% EPL master thesis covers template
\documentclass{EPL-master-thesis-covers-EN}

% Please fill in the following boxes
% Title of the thesis
\title{Benchmarking of Real-Time Operating Systems for Internet of Things}

% Subtitle - remove this line if not applicable
% \subtitle{Optional subtitle}

% Name of the student author(s)
\author{Julien \textsc{Gomez}}
\secondauthor{Trong-Vu \textsc{Tran}}		% remove if not applicable

\degreetitle{Master [120] in Computer Science}

% Name of the supervisor(s)
\supervisor{Ramin \textsc{Sadre}}

% Name of the reader(s)
\readerone{Lionel \textsc{Metongnon}}
\readertwo{Ludovic \textsc{Moreau}}

% Academic year (update if necessary)
\years{2018--2019}
%\documentclass{eplmastersthesis}
\usepackage{biblatex}
\usepackage{hyperref}
%\usepackage{pslatex}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{chngcntr}%%reset chapter count
\counterwithin*{chapter}{part}

% Listing package
\usepackage{listings}
\usepackage{xcolor}

\definecolor{mGreen}{rgb}{0,0.6,0}
\definecolor{mGray}{rgb}{0.5,0.5,0.5}
\definecolor{mPurple}{rgb}{0.58,0,0.82}

\lstdefinestyle{CStyle}{  
    commentstyle=\color{mGreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{mGray},
    stringstyle=\color{mPurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2,
    language=C
}

\lstdefinestyle{ascii-tree}{
  literate={├}{|}1 {─}{--}1 {└}{+}1,
  captionpos=b,          
}

\addbibresource{bibliography.bib}
\begin{document}

% Front cover page
  \maketitle

\cleardoublepage
\pagenumbering{roman}

%%%%%%% Abstract %%%%%
 \begin{abstract}
    % Here goes the abstract
    %1. What the thesis is about
    %2. The problem & the main argument
    %3. What you did (methods or techniques adopted, data collection and analysis of data)
    %4. findings & results
    %5. conclusion & future challenges (if possible)

    %Introduction
    The purpose of this thesis is to assess the possibilty to create a benchmarking framework aimed at 
      real-time operating systems running on constrained devices. 
    This benchmarking framework will be used to evaluate the performances of an application on top of different RTOS.
    The performance metric used is the context switching time.

    %Methods
    To do so, we started with a theoretical analysis of real-time operating systems.
    This allows us to formalize the concepts needed for the rest of the thesis.
    After that, we compared different RTOS in term of features and characteristics.
    It helped us understand and explain the operations we performed and the results we obtained.
    For the developpement of the benchmarking framework, we explored three different approaches which are described in separate sections.
    
    %Results
    Our results showed that using the internal real-time clock to compute the context switching time is not suited for the benchmarking framework.
    Instead, the use of an external board such as the Pocket Science Lab is the best option to retrieve precise measurements.

    %Discussion
    Future improvements include the use of a cache inside the RTOS to retrieve interessting data such as the memory usage or the tasks utilization.
    Moreover, the external board can be used to monitor the power consumption of the benchmarked board.
    
 \end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
 % Ramin Sadre (promoter)
 % Ludovic Moreau (lecturer)
 % Lionel Metongnon (lecturer)
 % Pascal Simon (welcome lab)
 % RIOT community
 % Contiki community
 % FreeRTOS community
 % Apache Mynewt community
 % mBed community
 % Emmanuel Baccelli
 % Thomas C. Schmidt 
 % Daniele Lacamera
 % Rahul Arvind Jadhav
 % Rodrigo Peixoto
 % Vicente J.P. Amorim
 % Jukka Rissanen
 % Victor Hansen
 % Salle Parnas
\end{abstract}

\renewcommand{\contentsname}{Content} %change contents to content

%%%%%%%% Table of contents %%%%%%
{\hypersetup{hidelinks=true}
% or \hypersetup{linkcolor=black}, if the colorlinks=true option of hyperref is used
\tableofcontents
}

\cleardoublepage
\pagenumbering{arabic}

\include{introduction}

\part{Comparing RTOS from a theoretical point of view}
%%%%%% Chap 1 %%%
\include{theory/index}

%%%%%% Chap 2 %%%
\include{rtos_outline/index}

\part{Implementing a benchmarking framework in the context of a RTOS}
%%%%%% Chap 3 %%%
\include{objective/index}

%%%%%% Chap 4 %%%
\include{experiment/index}

%%%%%% Chap 5 %%%
\include{results/index}

%%%%%% Chap 6 %%%
\include{discussion/index}


\include{conclusion}

\cleardoublepage
\pagenumbering{roman}
\listoffigures
\listoftables

%%%%%% Bibliography %%%
\printbibliography

% Back cover page
\backcoverpage
\end{document}