\part{Theory of RTOS\label{part:rtos-theory}}

\chapter{RTOS characteristics}

\paragraph{}
In order to understand how an RTOS works compared to a general purpose operating system,
    it is essential to define some reccurent characteristics we'd expect to find in a real-time architecture.
\\
This chapter is non exhaustive and one can debate about the importance of one relative to one other.
We decided to chose those characteristics because they are commonly used in the litterature
    and we think that they will provide a good glance at what we can expect from an RTOS.

The first part of this thesis is the result of an effort from ours to summarize the knowledge we collected during our researches.
The following informations do not come from a single source.
We spent countless hours collecting data, reading papers, documentation and technical datasheets from vendors, researchers and developpers.
We also spent time contacting qualified people to review, read and advise us on very specific topics.


\input{part1/characteristics/system_architecture}
\input{part1/characteristics/scheduling}
\input{part1/characteristics/memory_management}
\input{part1/characteristics/energy_management}
\input{part1/characteristics/programming_model}
\input{part1/characteristics/hardware_support}


\chapter{RTOS Outline}

\paragraph{}
In this chapter, we will cover the specificities of each operating system we chose to work with.
We chose to write about technical and non-technical characteristics of the chosen OS's and even their respectives projects as a whole.

Some of the things written in this chapter are personal opinions.
Indeed, as we decided to include non-technical characteristics of each of the projects, some subjectivity is involved.
We tried to be critical and impartial in those points but it beware it only reflects our own opinions.


A table summarizing the comparison is provided at the end of the chapter.

%https://www.simform.com/iot-rtos-selection/
%https://ijarcce.com/wp-content/uploads/2015/12/IJARCCE-92.pdf
%http://www.ipcsit.com/vol1/54-S028.pdf

%Supported Architecture
%License
%User-friendly/documentation/example applications
%Supported network protocols
%kernel type
%date of creation
%file systems supported
%number of boards supported
%memory footprint
%resource access control (posix...)
%memory management policy
%energy management features
%programming model
%scheduling
%community
%build system
%ipc management

\input{part1/rtos_outline/riot}
\input{part1/rtos_outline/contiki}
\input{part1/rtos_outline/comparison_table}